\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2022-02-21 3-advanced.Rmd
go through writing made yesterday
work on beginning of matching

Cubble provides functionality for matching sites from two data sources. To be a valid pair of matching, the matched pair from different data sources need to:

  - be spatially close to each other, and 
  - have similar temporal movement

\code{match_sites()} matches data based on these two criteria: it first matches the two data sources spatially through computing the pairwise distance on latitude and longitude. Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window. Figure \ref{fig:illu-matching} illustrates this temporal matching in more details. Given two series \code{A} and \code{a}, 3 peaks have been picked in each series. An interval, with default length of 5, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals. In this illustration, there are 2 matches for these two series. Several arguments are available in \code{match_sites()} to fine-tune the matching:

@@

This functionality can be seen as a matching exercise [@stuart2010matching; @mcintosh2018using] in the spatio-temporal domain to pair up nearby observations. It can also be considered as a form of data fusion [@castanedo2013review; @cocchi2019data], where data from multiple sources are combined together.

@@

## Spatio-temporal transformations

>>>>

One task that may interest analysts in spatio-temporal data is to find how similar the time series from nearby sites are. This can be seen as a matching problem [@stuart2010matching; @mcintosh2018using] that pairs up similar similar time series in nearby locations or a data fusing exercise that merges data collected from different sources [@cocchi2019data]. \code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task. The algorithm first matches the two data sources spatially through computing the pairwise distance on latitude and longitude. Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window. Figure \ref{fig:illu-matching} illustrates this temporal matching in more details. Given two series \code{A} and \code{a}, 3 peaks have been picked in each series. An interval, with default length of 5, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals. In this illustration, there are 2 matches for these two series. Several arguments are available in \code{match_sites()} to fine-tune the matching:

@@

@@

## Spatio-temporal transformations {#st_transformation}

### 2022-02-21 main.tex
work on beginning of matching

\textcolor{red}{more work on groupby} The grouping pair, \code{group_by}
and \code{ungroup}, can only be used in the long form to define grouping
on the time but not in the nested form.

@@

\textcolor{red}{more beginning} Cubble provides functionality for
matching sites from two data sources. To be a valid pair of matching,
the matched pair from different data sources need to:

\begin{itemize}
\tightlist
\item
  be spatially close to each other, and
\item
  have similar temporal movement
\end{itemize}

\code{match_sites()} matches data based on these two criteria: it first
matches the two data sources spatially through computing the pairwise
distance on latitude and longitude. Pairs that pass the spatial matching
are then matched temporally through computing the number of matched
peaks within a fixed length window. Figure \ref{fig:illu-matching}
illustrates this temporal matching in more details. Given two series
\code{A} and \code{a}, 3 peaks have been picked in each series. An
interval, with default length of 5, is constructed for each peak in
series \code{A} and the peaks in series \code{a} are tested against
whether they fall into the any of the intervals. In this illustration,
there are 2 matches for these two series. Several arguments are
available in \code{match_sites()} to fine-tune the matching:

@@

This functionality can be seen as a matching exercise
\citep{stuart2010matching, mcintosh2018using} in the spatio-temporal
domain to pair up nearby observations. It can also be considered as a
form of data fusion \citep{cocchi2019data}, where data from multiple
sources are combined together.

>>>>

@@

One task that may interest analysts in spatio-temporal data is to find
how similar the time series from nearby sites are. This can be seen as a
matching problem \citep{stuart2010matching, mcintosh2018using} that
pairs up similar similar time series in nearby locations or a data
fusing exercise that merges data collected from different sources
\citep{cocchi2019data}. \code{match_sites()} in \pkg{cubble} provides a
simple algorithm for this task. The algorithm first matches the two data
sources spatially through computing the pairwise distance on latitude
and longitude. Pairs that pass the spatial matching are then matched
temporally through computing the number of matched peaks within a fixed
length window. Figure \ref{fig:illu-matching} illustrates this temporal
matching in more details. Given two series \code{A} and \code{a}, 3
peaks have been picked in each series. An interval, with default length
of 5, is constructed for each peak in series \code{A} and the peaks in
series \code{a} are tested against whether they fall into the any of the
intervals. In this illustration, there are 2 matches for these two
series. Several arguments are available in \code{match_sites()} to
fine-tune the matching:

@@
